Assessing Gender Bias in Machine Translation -- A Case Study with Google Translate

### 2018-04-20 Paper/main.tex
Added paper
Major changes

\usepackage{multirow}
\usepackage{float}
\usepackage{array}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{pifont}

@@

We shall assume and then show that the phenomenon of gender bias in machine translation can be assessed by mapping sentences constructed in gender neutral languages to English by the means of an automated translation tool. Specifically, we can translate sentences such as the Hungarian ``ő ápolónő'', where ``ápolónő'' translates to ``nurse'' and ``ő'' is a gender-neutral pronoun meaning either he, she or it, to English, yielding in this example the result ``she's a nurse'' on Google Translate. As Figure \ref{fig:screenshot-gtranslate-hungarian} clearly shows, the same template yields a male pronoun when ``nurse'' is replaced by ``engineer''. The same basic template can be ported to all other gender neutral languages, as depicted in Table \ref{tab:templates}. Given the success of Google Translate, which amounts to 200 million users daily, we have chosen to exploit its API to obtain the desired thermometer of gender bias. Also, in order to solidify our results, we have decided to work with a fair amount of gender neutral languages, forming a list of these with help from the World Atlas of Language Structures (WALS) \citep{wals} and other sources. Table \ref{tab:gender-neutral-languages} compiles all languages we chose to use, with additional columns informing whether they
(1) exhibit a pronominal gender system and (2) are supported by Google Translate. Because pronominal gender systems defy the purposes of our technique, such languages have been discarded.

There is a prohibitively large class of nouns and adjectives that could in principle be substituted in the templates of Table \ref{tab:templates}. To simplify our dataset, we have decided to obtain a comprehensive list of professional occupations, which, we believe, are an interesting window into the nature of gender bias. Here, we resorted to using the Bureau of Labor Statistics' detailed occupations table \citep{BLS2017}, from the United States Department of Labor, for providing such a list. The values inside, however, had to be expanded since each line contained multiple occupations and sometimes very specific ones. Fortunately this table also provided a percentage of women participation in the jobs shown, for those that had more than 50 thousand workers. We filtered some of these because they were too generic ( ``Computer occupations, all other'', and others) or because they had gender specific words for the profession (``host/hostess'', ``waiter/waitress''). We then separated the curated jobs into broader categories (Artistic, Corporate, Theatre, etc.) as shown in Table \ref{tab:occupations}. Finally, Table \ref{tab:occupations-examples} shows thirty examples of randomly selected occupations from our dataset. For the occupations that had less than 50 thousand workers, and thus no data about the participation of women, we assumed that its women participation was that of its upper category. Finally, we have selected a small list of 21 adjectives, presented in Table \ref{tab:adjectives}. All code and data used to generate and compile the results presented in the following sections will be made publicly available in a Github repository upon the Editors request. They are not linked to external sites to avoid authors identification.

% Table containing a list of gender neutral languages with information about pronominal gender systems and GTranslate support
\begin{table}[H]
\centering
\begin{small}
	\begin{tabular}{|c|m{1.5cm}|m{1.5cm}|m{2.2cm}|m{1.5cm}|}
	\hline
	Language Family 				& Language & Supported 	& Phrases have male/female markers & Tested \\ \hline \hline
	\multirow{2}{*}{Austronesian} 	& Malay 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Tagalog 				& \xmark	 	& - 			& \xmark		\\ \hline
	\multirow{3}{*}{Finno-Ugric} 	& Estonian 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Finnish 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Hungarian 			& \checkmark 	& \xmark 		& \checkmark	\\ \hline
	\multirow{4}{*}{Indo-European} 	& Armenian 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Bengali				& \checkmark 	& O 			& \checkmark	\\
									& English 				& \checkmark 	& \checkmark 	& \xmark 		\\
									& {\color{red}Persian}	& \checkmark 	& \xmark 		& \checkmark	\\ \hline
	\multirow{3}{*}{Indo-Aryan} 	& Maithili 				& \xmark 	 	& - 			& \xmark 		\\
									& {\color{red}Nepali}	& \checkmark 	& O 	 		& \checkmark	\\
									& Oriya 				& \xmark 	 	& - 			& \xmark 		\\ \hline
	\multirow{10}{*}{} 				& Japanese 				& \checkmark 	& \xmark 		& \checkmark	\\
									& {\color{red}Korean}	& \checkmark 	& \checkmark	& \xmark		\\
									& Turkish 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Yoruba 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Basque 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Swahili 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Chinese 				& \checkmark 	& O 			& \checkmark	\\
									& Cantonese 			& \xmark	 	& - 			& \xmark		\\
									& Pipil 				& \xmark	 	& - 			& \xmark		\\
									& Quechuan 				& \xmark	 	& - 			& \xmark		\\ \hline
	\multirow{4}{*}{Constructed} 	& Esperanto 			& \checkmark 	& \checkmark 	& \xmark		\\
									& Ido 					& \xmark 	 	& - 			& \xmark 		\\ \cline{2-2}
									& Lingua Franca Nova 	& \xmark 	 	& - 			& \xmark		\\ \cline{2-2}
									& Interlingua 			& \xmark 	 	& - 			& \xmark		\\ \hline
	\end{tabular}
    \end{small}
	\caption{\changess{Selected gender neutral languages obtained from WALS and other sources. Languages are grouped according to language families and classified according to whether enforce any kind of mandatory gender (male/female) demarcation on simple phrases (\checkmark: yes, \xmark:~no, O: it is optional). For the purposes of this work, we have decided to work only with languages lacking such demarcation and as such Koran and Esperanto have been discarded. Languages lacking support from Google Translate have been discarded. Languages colored in red have been omitted in other reasons. See Section} \ref{ssec:rationale_for_language_exceptions} \changess{for further explanation.}}
	\label{tab:gender-neutral-languages}
\end{table}

@@

% List of sentence templates for assessing gender bias in the translation from source language to English
\begin{table}[H]
\centering
\small{
	\begin{tabular}{|c|c|}
	\hline
	Language 	& Sentence template 													\\ \hline \hline
	Malay		& dia adalah $\langle occupation \rangle$ 								\\ \hline
	Estonian	& ta on $\langle occupation \rangle$ 									\\ \hline
	Finnish		& hän on $\langle occupation \rangle$ 									\\ \hline
	Hungarian	& ő egy $\langle occupation \rangle$ 									\\ \hline
	Armenian	& na $\langle occupation \rangle$ e 									\\ \hline
	\multirow{6}{*}{Bengali}	& Ē ē ēkajana $\langle occupation \rangle$				\\
								& Yini yini ēkajana $\langle occupation \rangle$		\\
								& Ō ō ēkajana $\langle occupation \rangle$				\\
								& Uni uni ēkajana $\langle occupation \rangle$			\\
								& Sē sē ēkajana $\langle occupation \rangle$			\\
								& Tini tini ēkajana $\langle occupation \rangle$		\\ \hline
	Japanese	& $\langle occupation \rangle$ \begin{CJK*}{UTF8}{min}です\end{CJK*}		\\ \hline
	Turkish		& o bir $\langle occupation \rangle$ 									\\ \hline
	Yoruba		& o jẹ $\langle occupation \rangle$ 									\\ \hline
	Basque		& $\langle occupation \rangle$ da 										\\ \hline
	Swahili		& yeye ni $\langle occupation \rangle$ 									\\ \hline
	Chinese		& ta $\langle occupation \rangle$ 										\\ \hline
	\end{tabular}
	}
	\caption{Templates used to infer gender biases in the translation to the English language.}
	\label{tab:templates}
\end{table}

@@

% Table with all adjectives we experimented with
\begin{table}[H]
\centering
\small{
	\begin{tabular}{|c|c|c|}
	\hline
	Happy 		& Sad 		 & Right 	\\ \hline
	Wrong 		& Afraid	 & Brave 	\\ \hline
	Smart		& Dumb		 & Proud 	\\ \hline
	STRONG		& POLITE	 & Cruel 	\\ \hline
	DESIRABLE	& Loving	 & SYMPHATETIC 	\\ \hline
	Modest 		& SUCCESSFUL & Guilty	\\ \hline
	Innocent	& MATURE	 & Shy		\\ \hline
	\end{tabular}
	}
    \caption{Adjectives}
	\label{tab:adjectives}
\end{table}

@@

% PEDRO: TODO highlight table
% Table with our gender bias results grouped by category
\begin{table}[H]
\centering
\small{
	\begin{tabular}{|c|c|c|c|}
	\hline
	Category 								& Female ($\%$)	& Male ($\%$)	& Neutral ($\%$) \\ \hline
	\hline		
	Education / Training / Library 			& $23.737$ 		& $58.081$		& $3.535$		\\ \hline
	Business / Financial 					& $6.039$		& $62.319$		& $8.937$		\\ \hline
	Office / Administrative					& $9.962$		& $56.322$		& $9.451$		\\ \hline
	Healthcare support						& $27.083$		& $50.000$		& $9.722$		\\ \hline
	Management								& $11.594$		& $61.111$		& $4.831$		\\ \hline
	Installation / Maintenance / Repair		& $3.297$		& $62.393$		& $14.164$		\\ \hline
	Healthcare practitioners / technical	& $18.088$		& $61.499$		& $8.527$		\\ \hline
	Community and social service			& $14.286$ 		& $60.317$		& $5.556$		\\ \hline
	Sales and related						& $7.540$		& $64.683$		& $10.714$		\\ \hline
	Production								& $15.32$		& $45.370$		& $15.825$		\\ \hline
	Architecture and engineering			& $1.916$		& $75.862$		& $6.130$		\\ \hline
	Life, physical, and social sciences		& $3.595$		& $80.065$		& $4.575$		\\ \hline
	Transportation / material moving		& $7.778$		& $60.159$		& $14.286$		\\ \hline
	Arts / Entertainment					& $7.808$		& $66.366$		& $5.405$		\\ \hline
	Legal									& $9.524$ 		& $65.079$		& $6.349$		\\ \hline
	Protective service						& $6.746$		& $71.032$		& $7.143$		\\ \hline
	Food preparation / serving			& $14.379$ 		& $56.863$		& $12.418$		\\ \hline
	Farming, fishing, and forestry			& $12.821$ 		& $53.846$		& $10.256$		\\ \hline
	Computer and mathematical				& $3.472$		& $70.833$		& $5.556$		\\ \hline
	Personal care and service				& $17.845$ 		& $52.862$		& $11.111$		\\ \hline
	Construction and extraction				& $9.967$		& $54.739$		& $13.562$		\\ \hline
	Cleaning / Maintenance					& $12.222$		& $57.778$		& $10.000$		\\ \hline
	\hline
	Total           						& $11.078$ 		& $57.398$ 		& $11.286$ 		\\ \hline
	\end{tabular}
	}
    \caption{Percentage of female, male and neutral gender pronouns obtained for each BLS occupation category, averaged over all occupations in said category and tested languages detailed in Table \ref{tab:gender-neutral-languages}. Note that columns do not in general add up to $100\%$, as there is a fair amount of translated sentences for which we cannot obtain a gender pronoun.}
	\label{tab:gender-by-category}
\end{table}

% Table with our gender bias results grouped by category (but categories are further grouped into "super-categories")
\begin{table}[H]
\centering
\small{
	\begin{tabular}{|c|c|c|c|}
	\hline
	Category 								& Female ($\%$)			& Male ($\%$)			& Neutral ($\%$)	\\ \hline
	\hline
	Education								& $23.737$				& $58.081$				& $3.535$			\\ \hline
	Corporate								& $8.519$				& $62.407$				& $7.778$			\\ \hline
	Service									& $8.730$				& $59.746$				& $11.619$			\\ \hline
	Healthcare								& $20.527$				& $58.380$				& $8.851$			\\ \hline
	Production								& $15.32$				& $45.370$				& $15.825$			\\ \hline
	STEM									& $2.954$				& $76.653$				& $5.345$			\\ \hline
	Arts / Entertainment					& $7.808$				& $66.366$				& $5.405$			\\ \hline
	Legal									& $9.524$				& $65.079$				& $6.349$			\\ \hline
	Farming / Fishing / Forestry			& $12.821$				& $53.846$				& $10.256$			\\ \hline
	Construction / Extraction				& $9.967$				& $54.739$				& $13.562$			\\ \hline
	\hline
	Total           						& $11.078$ 				& $57.398$ 				& $11.286$ 			\\ \hline
	\end{tabular}
	}
    \caption{Percentage of female, male and neutral gender pronouns obtained for each of the merged occupation category, averaged over all occupations in said category and tested languages detailed in Table \ref{tab:gender-neutral-languages}. Note that columns do not in general add up to $100\%$, as there is a fair amount of translated sentences for which we cannot obtain a gender pronoun.}
	\label{tab:gender-by-category-grouped}
\end{table}

@@

% Histogram with the distribution of female pronouns for each occupation category
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histogram-female}
	\caption{The data for the total number $X$ of female pronouns per occupation category suggests an inverse distribution. STEM fields such as life, social, computer and physical sciences, architecture, engineering and mathematics are nearly exclusively concentrated at $X=0$, while more evenly distributed fields such as production and healthcare (see Table \ref{tab:gender-by-category}) extend to higher values.}
	\label{fig:histogram-female}
\end{figure}

% Histogram with the distribution of female pronouns for each occupation category (but categories are grouped into 'super-categories')
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histogram-female-grouped}
	\caption{Coalescing the data from Table \ref{tab:gender-by-category} by grouping categories into broader groups helps us see how translated gender pronouns are distributed for some fields of interest such as STEM, Education, Arts / Entertainment, etc. In this context it becomes even clearer that STEM fields concentrate at $X = 0$, as discussed in Figure \ref{fig:histogram-female}.}
	\label{fig:histogram-female-grouped}
\end{figure}

% Histogram with the distribution of male pronouns for each occupation category
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histogram-male}
	\caption{In contrast to female pronouns, male pronouns are seemingly skew normally distributed, with a peak at $X = 6$. One can see that STEM fields concentrate mainly to the right ($X \geq 6$).}
	\label{fig:histogram-male}
\end{figure}

% Histogram with the distribution of male pronouns for each occupation category (but categories are grouped into 'super-categories')
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histogram-male-grouped}
	\caption{Coalescing the data from Table \ref{tab:gender-by-category} by grouping categories into broader groups helps us see how translated gender pronouns are distributed for some fields of interest such as STEM, Education, Arts / Entertainment, etc. In this context it becomes clearer that STEM fields concentrate mainly to the right ($X \geq 6$), in contrast to what we observe for female pronouns (see Figure \ref{fig:histogram-female-grouped}).}
	\label{fig:histogram-male-grouped}
\end{figure}

% Histogram with the distribution of neutral pronouns for each occupation category
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histogram-neutral}
	\caption{The scarcity of gender-neutral pronouns is manifest in their histogram, showing how they exceed the rarity even of female pronouns (see Figure \ref{fig:histogram-female}). Once again, STEM fields are predominantly concentrated on $X = 0$.}
	\label{fig:histogram-neutral}
\end{figure}

% Histogram with the distribution of neutral pronouns for each occupation category (but categories are grouped into 'super-categories')
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histogram-neutral-grouped}
	\caption{Coalescing the data from Table \ref{tab:gender-by-category} by grouping categories into broader groups helps us see how translated gender pronouns are distributed for some fields of interest such as STEM, Education, Arts / Entertainment, etc. Once again, STEM fields are predominantly concentrated at $X = 0$.}
	\label{fig:histogram-neutral-grouped}
\end{figure}

>>>>

\usepackage{multirow}
\usepackage{float}
\usepackage{array}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage[caption=false]{subfig}

@@

We shall assume and then show that the phenomenon of gender bias in machine translation can be assessed by mapping sentences constructed in gender neutral languages to English by the means of an automated translation tool. Specifically, we can translate sentences such as the Hungarian ``ő ápolónő'', where ``ápolónő'' translates to ``nurse'' and ``ő'' is a gender-neutral pronoun meaning either he, she or it, to English, yielding in this example the result ``she's a nurse'' on Google Translate. As Figure \ref{fig:screenshot-gtranslate-hungarian} clearly shows, the same template yields a male pronoun when ``nurse'' is replaced by ``engineer''. The same basic template can be ported to all other gender neutral languages, as depicted in Table \ref{tab:templates-occupations}. Given the success of Google Translate, which amounts to 200 million users daily, we have chosen to exploit its API to obtain the desired thermometer of gender bias. Also, in order to solidify our results, we have decided to work with a fair amount of gender neutral languages, forming a list of these with help from the World Atlas of Language Structures (WALS) \citep{wals} and other sources. Table \ref{tab:gender-neutral-languages} compiles all languages we chose to use, with additional columns informing whether they
(1) exhibit a pronominal gender system and (2) are supported by Google Translate. Because pronominal gender systems defy the purposes of our technique, such languages have been discarded.

\changess{There is a prohibitively large class of nouns and adjectives that could in principle be substituted into our templates. To simplify our dataset, we have decided to focus our work on job positions -- which, we believe, are an interesting window into the nature of gender bias --, and were able to obtain a comprehensive list of professional occupations from the Bureau of Labor Statistics' detailed occupations table} \citep{BLS2017}\changess{, from the United States Department of Labor.} The values inside, however, had to be expanded since each line contained multiple occupations and sometimes very specific ones. Fortunately this table also provided a percentage of women participation in the jobs shown, for those that had more than 50 thousand workers. We filtered some of these because they were too generic ( ``Computer occupations, all other'', and others) or because they had gender specific words for the profession (``host/hostess'', ``waiter/waitress''). We then separated the curated jobs into broader categories (Artistic, Corporate, Theatre, etc.) as shown in Table \ref{tab:occupations}. Finally, Table \ref{tab:occupations-examples} shows thirty examples of randomly selected occupations from our dataset. For the occupations that had less than 50 thousand workers, and thus no data about the participation of women, we assumed that its women participation was that of its upper category. \changess{Finally, as complementary evidence we have decided to include a small subset of 21 adjectives in our study. All adjectives were obtained from the top one thousand most frequent words in this category as featured in the Corpus of Contemporary American English (COCA)} \url{https://corpus.byu.edu/coca/}\changess{, but it was necessary to manually curate them because a substantial fraction of these adjectives cannot be applied to human subjects. Also because the sentiment associated with each adjective is not as easily accessible as for example the occupation category of each job position, we performed a manual selection of a subset of such words which we believe to be meaningful to this study. These words are presented in Table} \ref{tab:adjectives}. All code and data used to generate and{} compile the results presented in the following sections will be made publicly available in a Github repository upon the Editors request. They are not linked to external sites to avoid authors identification.

% Table containing a list of gender neutral languages with information about pronominal gender systems and GTranslate support
\begin{table}[H]
\centering
\begin{small}
	\begin{tabular}{|c|m{1.5cm}|m{1.5cm}|m{2.2cm}|m{1.5cm}|}
	\hline
	Language Family 				& Language & Supported 	& Phrases have male/female markers & Tested \\ \hline \hline
	\multirow{2}{*}{Austronesian} 	& Malay 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Tagalog 				& \xmark	 	& - 			& \xmark		\\ \hline
	\multirow{3}{*}{Finno-Ugric} 	& Estonian 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Finnish 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Hungarian 			& \checkmark 	& \xmark 		& \checkmark	\\ \hline
	\multirow{4}{*}{Indo-European} 	& Armenian 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Bengali				& \checkmark 	& O 			& \checkmark	\\
									& English 				& \checkmark 	& \checkmark 	& \xmark 		\\
									& {\color{red}Persian}	& \checkmark 	& \xmark 		& \checkmark	\\ \hline
	\multirow{3}{*}{Indo-Aryan} 	& Maithili 				& \xmark 	 	& - 			& \xmark 		\\
									& {\color{red}Nepali}	& \checkmark 	& O 	 		& \checkmark	\\
									& Oriya 				& \xmark 	 	& - 			& \xmark 		\\ \hline
	\multirow{10}{*}{} 				& Japanese 				& \checkmark 	& \xmark 		& \checkmark	\\
									& {\color{red}Korean}	& \checkmark 	& \checkmark	& \xmark		\\
									& Turkish 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Yoruba 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Basque 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Swahili 				& \checkmark 	& \xmark 		& \checkmark	\\
									& Chinese 				& \checkmark 	& O 			& \checkmark	\\
									& Cantonese 			& \xmark	 	& - 			& \xmark		\\
									& Pipil 				& \xmark	 	& - 			& \xmark		\\
									& Quechuan 				& \xmark	 	& - 			& \xmark		\\ \hline
	\multirow{4}{*}{Constructed} 	& Esperanto 			& \checkmark 	& \checkmark 	& \xmark		\\
									& Ido 					& \xmark 	 	& - 			& \xmark 		\\ \cline{2-2}
									& Lingua Franca Nova 	& \xmark 	 	& - 			& \xmark		\\ \cline{2-2}
									& Interlingua 			& \xmark 	 	& - 			& \xmark		\\ \hline
	\end{tabular}
    \end{small}
	\caption{\changess{Selected gender neutral languages obtained from WALS and other sources. Languages are grouped according to language families and classified according to whether enforce any kind of mandatory gender (male/female) demarcation on simple phrases (\checkmark: yes, \xmark:~no, O: it is optional). For the purposes of this work, we have decided to work only with languages lacking such demarcation such as Esperanto have been discarded. Languages lacking support from Google Translate have been discarded. Languages colored in red have been omitted in other reasons. See Section} \ref{ssec:rationale_for_language_exceptions} \changess{for further explanation.}}
	\label{tab:gender-neutral-languages}
\end{table}

@@

% List of sentence templates for assessing gender bias in the translation from source language to English
\begin{table}[H]
\centering
\small{
	\begin{tabular}{|c|c|}
	\hline
	Language 	& Sentence template 													\\ \hline \hline
	Malay		& dia adalah $\langle occupation \rangle$ 								\\ \hline
	Estonian	& ta on $\langle occupation \rangle$ 									\\ \hline
	Finnish		& hän on $\langle occupation \rangle$ 									\\ \hline
	Hungarian	& ő egy $\langle occupation \rangle$ 									\\ \hline
	Armenian	& na $\langle occupation \rangle$ e 									\\ \hline
	\multirow{6}{*}{Bengali}	& Ē ēkajana $\langle occupation \rangle$				\\
								& Yini ēkajana $\langle occupation \rangle$		\\
								& Ō ēkajana $\langle occupation \rangle$				\\
								& Uni ēkajana $\langle occupation \rangle$			\\
								& Sē ēkajana $\langle occupation \rangle$			\\
								& Tini ēkajana $\langle occupation \rangle$		\\ \hline
	Japanese	& \begin{CJK*}{UTF8}{min}あの人は\end{CJK*} $\langle occupation \rangle$ \begin{CJK*}{UTF8}{min}です\end{CJK*}		\\ \hline
	Turkish		& o bir $\langle occupation \rangle$ 									\\ \hline
	Yoruba		& o jẹ $\langle occupation \rangle$ 									\\ \hline
	Basque		& $\langle occupation \rangle$ bat da 										\\ \hline
	Swahili		& yeye ni $\langle occupation \rangle$ 									\\ \hline
	Chinese		& ta shi $\langle occupation \rangle$ 										\\ \hline
	\end{tabular}
	}
	\caption{\changess{Templates used to infer gender biases in the translation of job occupations to the English language.}}
	\label{tab:templates-occupations}
\end{table}

\begin{table}[H]
\centering
\small{
	\begin{tabular}{|c|c|}
	\hline
	Language 	& Sentence template 													\\ \hline \hline
	Malay		& dia $\langle adjective \rangle$ 								\\ \hline
	Estonian	& ta on $\langle adjective \rangle$ 									\\ \hline
	Finnish		& hän on $\langle adjective \rangle$ 									\\ \hline
	Hungarian	& ő $\langle adjective \rangle$ 									\\ \hline
	Armenian	& na $\langle adjective \rangle$ e 									\\ \hline
	\multirow{6}{*}{Bengali}	& Ē $\langle adjective \rangle$				\\
								& Yini $\langle adjective \rangle$		\\
								& Ō $\langle adjective \rangle$				\\
								& Uni $\langle adjective \rangle$			\\
								& Sē $\langle adjective \rangle$			\\
								& Tini $\langle adjective \rangle$		\\ \hline
	Japanese	& \begin{CJK*}{UTF8}{min}あの人は\end{CJK*} $\langle adjective \rangle$ \begin{CJK*}{UTF8}{min}です\end{CJK*}		\\ \hline
	Turkish		& o $\langle adjective \rangle$ 									\\ \hline
	Yoruba		& o jẹ $\langle adjective \rangle$ 									\\ \hline
	Basque		& $\langle adjective \rangle$ da 										\\ \hline
	Swahili		& yeye ni $\langle adjective \rangle$ 									\\ \hline
	Chinese		& ta hen $\langle adjective \rangle$ 										\\ \hline
	\end{tabular}
	}
	\caption{\changess{Templates used to infer gender biases in the translation of adjectives to the English language.}}
	\label{tab:templates-adjectives}
\end{table}

@@

% Table with all adjectives we experimented with
\begin{table}[H]
\centering
\small{
	\begin{tabular}{|c|c|c|}
	\hline
	Happy 		& Sad 		 & Right 		\\ \hline
	Wrong 		& Afraid	 & Brave 		\\ \hline
	Smart		& Dumb		 & Proud 		\\ \hline
	Strong		& Polite	 & Cruel 		\\ \hline
	Desirable	& Loving	 & Sympathetic 	\\ \hline
	Modest 		& Successful & Guilty		\\ \hline
	Innocent	& Mature	 & Shy			\\ \hline
	\end{tabular}
	}
    \caption{\changess{Curated list of 21 adjectives obtained from the top one thousand most frequent words in this category in the Corpus of Contemporary American English (COCA)} \url{https://corpus.byu.edu/coca/}.}
	\label{tab:adjectives}
\end{table}

@@

% PEDRO: TODO highlight table
% Table with our gender bias results grouped by category
\begin{table}[H]
\centering
\small{
	\begin{tabular}{|c|c|c|c|}
	\hline
	Category & Female ($\%$)	& Male ($\%$)	& Neutral ($\%$) \\ \hline
	\hline
	Office and administrative support & 11.015 & 58.812 & 16.954 \\ \hline
	Architecture and engineering & 2.299 & 72.701 & 10.92 \\ \hline
	Farming, fishing, and forestry & 12.179 & 62.179 & 14.744 \\ \hline
	Management & 11.232 & 66.667 & 12.681 \\ \hline
	Community and social service & 20.238 & 62.5 & 10.119 \\ \hline
	Healthcare support & 25.0 & 43.75 & 17.188 \\ \hline
	Sales and related & 8.929 & 62.202 & 16.964 \\ \hline
	Installation, maintenance, and repair & 5.22 & 58.333 & 17.125 \\ \hline
	Transportation and material moving & 8.81 & 62.976 & 17.5 \\ \hline
	Legal & 11.905 & 72.619 & 10.714 \\ \hline
	Business and financial operations & 7.065 & 67.935 & 15.58 \\ \hline
	Life, physical, and social science & 5.882 & 73.284 & 10.049 \\ \hline
	Arts, design, entertainment, sports, and media & 10.36 & 67.342 & 11.486 \\ \hline
	Education, training, and library & 23.485 & 53.03 & 9.091 \\ \hline
	Building and grounds cleaning and maintenance & 12.5 & 68.333 & 11.667 \\ \hline
	Personal care and service & 18.939 & 49.747 & 18.434 \\ \hline
	Healthcare practitioners and technical & 22.674 & 51.744 & 15.116 \\ \hline
	Production & 14.331 & 51.199 & 18.245 \\ \hline
	Computer and mathematical & 4.167 & 66.146 & 14.062 \\ \hline
	Construction and extraction & 8.578 & 61.887 & 17.525 \\ \hline
	Protective service & 8.631 & 65.179 & 12.5 \\ \hline
	Food preparation and serving related & 21.078 & 58.333 & 17.647 \\ \hline \hline
	Total & 11.76 & 58.93 & 15.939 \\ \hline

	\end{tabular}
	}
    \caption{\changess{Percentage of female, male and neutral gender pronouns obtained for each BLS occupation category, averaged over all occupations in said category and tested languages detailed in Table} \ref{tab:gender-neutral-languages}\changess{. Note that columns do not in general add up to $100\%$, as there is a fair amount of translated sentences for which we cannot obtain a gender pronoun.}}
	\label{tab:gender-by-category}
\end{table}

% Table with our gender bias results grouped by category (but categories are further grouped into "super-categories")
\begin{table}[H]
\centering
\small{
	\begin{tabular}{|c|c|c|c|}
	\hline
	Category 								& Female ($\%$)			& Male ($\%$)			& Neutral ($\%$)	\\ \hline
	\hline
	Service & 10.5 & 59.548 & 16.476 \\ \hline
	STEM & 4.219 & 71.624 & 11.181 \\ \hline
	Farming / Fishing / Forestry & 12.179 & 62.179 & 14.744 \\ \hline
	Corporate & 9.167 & 66.042 & 14.861 \\ \hline
	Healthcare & 23.305 & 49.576 & 15.537 \\ \hline
	Legal & 11.905 & 72.619 & 10.714 \\ \hline
	Arts / Entertainment & 10.36 & 67.342 & 11.486 \\ \hline
	Education & 23.485 & 53.03 & 9.091 \\ \hline
	Production & 14.331 & 51.199 & 18.245 \\ \hline
	Construction / Extraction & 8.578 & 61.887 & 17.525 \\ \hline \hline
	Total & 11.76 & 58.93 & 15.939 \\ \hline

	\end{tabular}
	}
    \caption{\changess{Percentage of female, male and neutral gender pronouns obtained for each of the merged occupation category, averaged over all occupations in said category and tested languages detailed in Table} \ref{tab:gender-neutral-languages}\changess{. Note that columns do not in general add up to $100\%$, as there is a fair amount of translated sentences for which we cannot obtain a gender pronoun.}}
	\label{tab:gender-by-category-grouped}
\end{table}

@@

% Histogram with the distribution of female pronouns for each occupation category
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histograms/all-categories-stacked-Female}
	\caption{\changess{The data for the total number $X$ of female pronouns per occupation category suggests an inverse distribution. STEM fields such as life, social, computer and physical sciences, architecture, engineering and mathematics are nearly exclusively concentrated at $X=0$, while more evenly distributed fields such as production and healthcare (see Table} \ref{tab:gender-by-category}\changess{) extend to higher values.}}
	\label{fig:histogram-female}
\end{figure}

% Histogram with the distribution of female pronouns for each occupation category (but categories are grouped into 'super-categories')
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histograms/all-categories-stacked-Female-grouped}
	\caption{\changess{Coalescing the data from Table} \ref{tab:gender-by-category} \changess{by grouping categories into broader groups helps us see how translated gender pronouns are distributed for some fields of interest such as STEM, Education, Arts / Entertainment, etc. In this context it becomes even clearer that STEM fields concentrate at $X = 0$, as discussed in Figure} \ref{fig:histogram-female}.}
	\label{fig:histogram-female-grouped}
\end{figure}

% Histogram with the distribution of male pronouns for each occupation category
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histograms/all-categories-stacked-Male}
	\caption{\changess{In contrast to female pronouns, male pronouns are seemingly skew normally distributed, with a peak at $X = 6$. One can see that STEM fields concentrate mainly to the right ($X \geq 6$).}}
	\label{fig:histogram-male}
\end{figure}

% Histogram with the distribution of male pronouns for each occupation category (but categories are grouped into 'super-categories')
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histograms/all-categories-stacked-Male-grouped}
	\caption{\changess{Coalescing the data from Table} \ref{tab:gender-by-category} \changess{by grouping categories into broader groups helps us see how translated gender pronouns are distributed for some fields of interest such as STEM, Education, Arts / Entertainment, etc. In this context it becomes clearer that STEM fields concentrate mainly to the right ($X \geq 6$), in contrast to what we observe for female pronouns (see Figure} \ref{fig:histogram-female-grouped}).}
	\label{fig:histogram-male-grouped}
\end{figure}

% Histogram with the distribution of neutral pronouns for each occupation category
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histograms/all-categories-stacked-Neutral}
	\caption{\changess{The scarcity of gender-neutral pronouns is manifest in their histogram, showing how they exceed the rarity even of female pronouns (see Figure} \ref{fig:histogram-female}\changess{). Once again, STEM fields are predominantly concentrated on $X = 0$.}}
	\label{fig:histogram-neutral}
\end{figure}

% Histogram with the distribution of neutral pronouns for each occupation category (but categories are grouped into 'super-categories')
\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histograms/all-categories-stacked-Neutral-grouped}
	\caption{\changess{Coalescing the data from Table} \ref{tab:gender-by-category} \changess{by grouping categories into broader groups helps us see how translated gender pronouns are distributed for some fields of interest such as STEM, Education, Arts / Entertainment, etc. Once again, STEM fields are predominantly concentrated at $X = 0$.}}
	\label{fig:histogram-neutral-grouped}
\end{figure}

AAAAAAAA

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histograms/categories/gender-dodged-STEM}
	\caption{TODO STEM}
	\label{fig:histogram-dodged-STEM}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/histograms/categories/gender-dodged-Healthcare}
	\caption{TODO Healthcare}
	\label{fig:histogram-dodged-Healthcare}
\end{figure}

BBBBBVBBV

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/heatmap-languages-categories-Female}
	\caption{TODO heatmap female}
	\label{fig:heatmap-female}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/heatmap-languages-categories-Male}
	\caption{TODO heatmap male}
	\label{fig:heatmap-male}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=10cm]{pictures/heatmap-languages-categories-Neutral}
	\caption{TODO heatmap neutral}
	\label{fig:heatmap-neutral}
\end{figure}
